---
title: "Assignment3"
output: word_document
date: "2022-10-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = list(CRAN="http://cran.rstudio.com/"))
require("knitr")
opts_knit$set(root.dir <- "~/Desktop/uf/FALL22/bioinformatics")
library(org.Hs.eg.db) 
library(magrittr)
install.packages("factoextra")
install.packages("arules")
library(stats)
```

#2
```{r}
library(magrittr)
library(dplyr) 
library(tibble)

compute_cv <- function(x) sd(x) / mean(x)
cv <- data.frame(apply(filtered_expression_df,1 , compute_cv))
colnames(cv) <- (c("cv"))
summary(cv)
#https://jdblischak.github.io/singlecell-qtl/pca-variable.html

orderedByCV <- data.frame(cv %>% arrange(desc(cv))) %>% rownames_to_column('gene')

#(a)
V5000 <- filtered_expression_df %>% filter(rownames(filtered_expression_df) %in% (orderedByCV[0:5000,1]))#5000 most variable genes
V5000 <- t(V5000)
```
```{r}
#(b)

##kmeans clustering
k2 <- kmeans(V5000, centers = 2)
k3 <- kmeans(V5000, centers = 3)
str(kmeans)
#https://uc-r.github.io/kmeans_clustering#prep
k4 <- kmeans(V5000, centers = 4)



library(factoextra)
fviz_cluster(k2, V5000)
fviz_cluster(k3, V5000)
fviz_cluster(k4, V5000)

##hierarchical clustering
library(stats)

hclust <- hclust(dist(V5000), method = "complete", members = NULL)
plot(hclust)

##consensus clustering 
BiocManager::install("ConsensusClusterPlus")
library(ConsensusClusterPlus)

v5000matrix <-data.matrix(V5000, rownames.force = NA)
results <- ConsensusClusterPlus(v5000matrix,maxK=4,reps=50,pItem=0.8,pFeature=1,title='title',clusterAlg="hc",distance="pearson",plot="screen")
```
##d
  Using kmeans and an initial cluster number of 2, one small cluster is nested within a cluster that was much larger in terms of area and membership. With k=3, the small nested cluster persists, but the larger cluster is split into two. As in the previous case, there is a cluster with a membership that is much larger than the other clusters. The same pattern can be oberved for k=4.

  The hierarchical clustering function doesn't appear to have chosen a k or been determined by any input for k. Instead it llooks like the clustering resolved down to each individual sample while providing a metric for cluster proximity with its height axis. In this case, the three or four primary clusters still maintain an extreme inequality in terms of sample membership, similar to our kmeans clustering results.

  Using consensus clustering, our delta area graph shows a cluster number of 3 as the most significant. The consensus clustering graph shows a majority of membership in one group with a smaller group and an even smaller group next to that. With an increase to a cluster number of 4, the appearance of a larger group emerges. This group is possibly analogous to the nested group emergin in the kmeans clustering with a k=4 as well.
  
In each of the methods, the proportion of cluster membership does not seem to reflect that of the control groups for our experiment, which is roughly half and half.


##(e) rerun each clustering method for 10, 100, 1000 and 10000 genes
```{r}
V10 <- filtered_expression_df %>% filter(rownames(filtered_expression_df) %in% (orderedByCV[0:10,1]))
V10 <- t(V10)
V100 <- filtered_expression_df %>% filter(rownames(filtered_expression_df) %in% (orderedByCV[0:100,1]))
V100 <- t(V100)
V1000 <- filtered_expression_df %>% filter(rownames(filtered_expression_df) %in% (orderedByCV[0:1000,1]))
V1000 <- t(V1000)
V10000 <- filtered_expression_df %>% filter(rownames(filtered_expression_df) %in% (orderedByCV[0:10000,1]))
V10000 <- t(V10000)
```
```{r}
##10
k2_10 <- kmeans(V10, centers = 2)
k3_10 <- kmeans(V10, centers = 3)
k4_10 <- kmeans(V10, centers = 4)

fviz_cluster(k2_10, V10)
fviz_cluster(k3_10, V10)
fviz_cluster(k4_10, V10)


hclust10 <- hclust(dist(V10), method = "complete", members = NULL)
plot(hclust10)

v10matrix <-data.matrix(V10, rownames.force = NA)
results10 <- ConsensusClusterPlus(v10matrix,maxK=4,reps=50,pItem=0.8,pFeature=1,title='title',clusterAlg="hc",distance="pearson",plot="screen")
```


```{r}
##100
k2_100 <- kmeans(V100, centers = 2)
k3_100 <- kmeans(V100, centers = 3)
k4_100 <- kmeans(V100, centers = 4)

fviz_cluster(k2_100, V100)
fviz_cluster(k3_100, V100)
fviz_cluster(k4_100, V100)

hclust100 <- hclust(dist(V100), method = "complete", members = NULL)
plot(hclust100)

v100matrix <-data.matrix(V100, rownames.force = NA)
results100 <- ConsensusClusterPlus(v100matrix,maxK=4,reps=50,pItem=0.8,pFeature=1,title='title',clusterAlg="hc",distance="pearson",plot="screen")
```

```{r}
#1000
k2_1000 <- kmeans(V1000, centers = 2)
k3_1000 <- kmeans(V1000, centers = 3)
k4_1000 <- kmeans(V1000, centers = 4)

fviz_cluster(k2_1000, V1000)
fviz_cluster(k3_1000, V1000)
fviz_cluster(k4_1000, V1000)

hclust1000 <- hclust(dist(V1000), method = "complete", members = NULL)
plot(hclust1000)

v1000matrix <-data.matrix(V1000, rownames.force = NA)
results1000 <- ConsensusClusterPlus(v1000matrix,maxK=4,reps=50,pItem=0.8,pFeature=1,title='title',clusterAlg="hc",distance="pearson",plot="screen")
```
```{r}
#10000
k2_10000 <- kmeans(V10000, centers = 2)
k3_10000 <- kmeans(V10000, centers = 3)
k4_10000 <- kmeans(V10000, centers = 4)

fviz_cluster(k2_10000, V10000)
fviz_cluster(k3_10000, V10000)
fviz_cluster(k4_10000, V10000)

hclust10000 <- hclust(dist(V10000), method = "complete", members = NULL)
plot(hclust10000)

v10000matrix <-data.matrix(V10000, rownames.force = NA)
results10000 <- ConsensusClusterPlus(v10000matrix,maxK=4,reps=50,pItem=0.8,pFeature=1,title='title',clusterAlg="hc",distance="pearson",plot="screen")
```




